Discrete Quantum Theories and Computing

### 2018-03-10 dissertation.tex
dissertation.lyx "2.2 Quantum Circuit Model" Updating...

\todo{Table of symbols?}

@@

An $n$-qubit pure state in the conventional quantum theory (CQT)
is represented as a vector in the $2^{n}$-dimensional Hilbert space.
If we eliminate all symmetries, an irreducible state is actually a
point in the projective Hilbert space \cite{MosseriDandoloff2001,Jaeger2007}
also known as the complex projective space~$\mathbb{CP}^{2^{n}-1}$
\cite{Hatcher2001,Bengtsson2007}. These irreducible quantum states
can also be classified as product states and entangled states, and
the latter one plays the essential role for pure-state algorithms
\cite{Mermin2007,Jaeger2007}.

@@

Entanglement may be regarded as one of the main characteristics distinguishing
quantum from classical mechanics. Entanglement involves quantum correlations
such that the measurement outcomes in one subsystem are related to
the measurement outcomes in another one. Within the standard framework,
given a quantum system composed of $n$ qubit subsystems, a pure state
of the total system $\ket{\Psi}$ is said to be entangled if it cannot
be written as a product of states of each subsystem. That is, a state
$\ket{\Psi}$ is entangled if 
\begin{eqnarray}
\ket{\Psi}\ne\ket{\psi_{1}}\otimes\cdots\otimes\ket{\psi_{j}}\otimes\cdots\otimes\ket{\psi_{n}},
\end{eqnarray}
where $\ket{\psi_{j}}$ refers to an arbitrary state of the $j$-th
qubit, and $\otimes$ represents the tensor product. This is equivalent
to saying that if one calculates the reduced density operator $\rho_{j}$
of the $j$th subsystem by tracing out all the other subsystems, $\rho_{j}={\sf tr}_{\{1,\cdots,j-1,j+1,\cdots,n\}}(\rho)$,
with $j=1,\cdots,n$ and $\rho=\ket{\Psi}\bra{\Psi}$, the normalized
state $\ket{\Psi}$, $\langle\Psi|\Psi\rangle=1$, is entangled if
and only if at least one subsystem state is \textit{mixed\/}; i.e.,
${\sf tr}_{j}(\rho_{j}^{2})<1$. For example, consider 
\begin{eqnarray}
\rho_{j}=\frac{1}{2}(\mathds{1}+\sum\limits _{\mu=x,y,z}\langle\sigma_{\mu}^{j}\rangle\ \sigma_{\mu}^{j}),
\end{eqnarray}
where $\sigma_{\mu}^{j}$, $\mu=x,y,z$, are the Pauli operators acting
on the $j$-th spin, 
\begin{equation}
\sigma_{\mu}^{j}=\overbrace{\mathds{1}\otimes\mathds{1}\otimes\cdots\otimes\underbrace{\sigma_{\mu}}_{j^{th}\ \mbox{factor}}\otimes\cdots\otimes\mathds{1}}^{n\ \mbox{factors}}\;,\label{pauli3}
\end{equation}
with 
\begin{eqnarray}
\hspace*{-0.5cm}\sigma_{x}=\left(\begin{array}{cc}
0 & 1\\
1 & 0
\end{array}\right),\sigma_{y}=\left(\begin{array}{cc}
0 & -i\\
i & 0
\end{array}\right),\sigma_{z}=\left(\begin{array}{cc}
1 & 0\\
0 & -1
\end{array}\right),
\end{eqnarray}
and $\langle\sigma_{\mu}^{j}\rangle=\langle\Psi|\sigma_{\mu}^{j}|\Psi\rangle$
denotes the corresponding expectation value. The vectors ${\bf X}_{j}=(\langle\sigma_{x}^{j}\rangle,\langle\sigma_{y}^{j}\rangle,\langle\sigma_{y}^{j}\rangle)$,
${\bf X}_{j}\in{\mathbb{R}}^{3}$, allow a geometric representation
of each reduced state in ${\mathbb{R}}^{3}$, satisfying $0\le\|{\bf X}_{j}\|\le1$.
Since ${\sf tr}_{j}(\rho_{j}^{2})=\frac{1}{2}(1+\|{\bf X}_{j}\|^{2})$,
the state $\ket{\Psi}$ is entangled if $\|{\bf X}_{j}\|<1$ for at
least one $j$, represented by a point {\em inside} the corresponding
local Bloch sphere. One may therefore consider $\ket{\Psi}$ to be
maximally entangled if $\|{\bf X}_{j}\|=0$ for all $j$. On the other
hand, the state $\ket{\Psi}$ is unentangled (i.e., a product state)
if $\|{\bf X}_{j}\|=1$ for all $j$, corresponding to points lying
on the surface of the Bloch sphere.

@@

>>>>

@@

\begin{comment}
An $n$-qubit pure state in the conventional quantum theory (CQT)
is represented as a vector in the $2^{n}$-dimensional Hilbert space.
If we eliminate all symmetries, an irreducible state is actually a
point in the projective Hilbert space \cite{MosseriDandoloff2001,Jaeger2007}
also known as the complex projective space~$\CP{2^{n}-1}$ \cite{Hatcher2001,Bengtsson2007}.
These irreducible quantum states can also be classified as product
states and entangled states, and the latter one plays the essential
role for pure-state algorithms \cite{Mermin2007,Jaeger2007}.
\end{comment}
It has been 36 years since Feynman~\cite{Feynman1982Simulating}
proposed the idea about quantum computation. Although there are many
models for quantum computing, the circuit model is still one of the
most widespread models~\cite{544199} no matter whether in theoretical
study or in building a real quantum computation. Information in a
quantum circuit is stored in qubits. Each qubit is a two-dimensional
quantum system. For a two-qubit system, if the first and second qubit
are described as the state~$\ket{\psi}=\alpha_{0}\ket{0}+\alpha_{1}\ket{1}$
and $\ket{\phi}=\beta_{0}\ket{0}+\beta_{1}\ket{1}$, the state of
the whole system is their tensor product 
\begin{equation}
\begin{aligned} & \ket{\psi}\ket{\phi}=\ket{\psi}\otimes\ket{\phi}=\left(\alpha_{0}\ket{0}+\alpha_{1}\ket{1}\right)\otimes\left(\beta_{0}\ket{0}+\beta_{1}\ket{1}\right)\\
= & \alpha_{0}\beta_{0}\ket{0}\otimes\ket{0}+\alpha_{0}\beta_{1}\ket{0}\otimes\ket{1}+\alpha_{1}\beta_{0}\ket{1}\otimes\ket{0}+\alpha_{1}\beta_{1}\ket{1}\otimes\ket{1}\\
= & \alpha_{0}\beta_{0}\ket{0}\ket{0}+\alpha_{0}\beta_{1}\ket{0}\ket{1}+\alpha_{1}\beta_{0}\ket{1}\ket{0}+\alpha_{1}\beta_{1}\ket{1}\ket{1}
\end{aligned}
\end{equation}
.

@@

Entanglement may be regarded as one of the main characteristics distinguishing
quantum from classical mechanics. Entanglement involves quantum correlations
such that the measurement outcomes in one subsystem are related to
the measurement outcomes in another one. Within the standard framework,
given a quantum system composed of $n$ qubit subsystems, a pure state
of the total system $\ket{\Psi}$ is said to be entangled if it cannot
be written as a product of states of each subsystem. That is, a state
$\ket{\Psi}$ is entangled if $\ket{\Psi}\ne\ket{\psi_{1}}\otimes\cdots\otimes\ket{\psi_{j}}\otimes\cdots\otimes\ket{\psi_{n}}$,
where $\ket{\psi_{j}}$ refers to an arbitrary state of the $j$-th
qubit, and $\otimes$ represents the tensor product. This is equivalent
to saying that if one calculates the reduced density operator $\rho_{j}$
of the $j$th subsystem by tracing out all the other subsystems, $\rho_{j}={\sf tr}_{\{1,\cdots,j-1,j+1,\cdots,n\}}(\rho)$,
with $j=1,\cdots,n$ and $\rho=\ket{\Psi}\bra{\Psi}$, the normalized
state $\ket{\Psi}$, $\langle\Psi|\Psi\rangle=1$, is entangled if
and only if at least one subsystem state is \textit{mixed\/}; i.e.,
${\sf tr}_{j}(\rho_{j}^{2})<1$. For example, consider 
\begin{eqnarray}
\rho_{j}=\frac{1}{2}(\mathds{1}+\sum\limits _{\mu=x,y,z}\langle\sigma_{\mu}^{j}\rangle\ \sigma_{\mu}^{j}),
\end{eqnarray}
where $\sigma_{\mu}^{j}$, $\mu=x,y,z$, are the Pauli operators acting
on the $j$-th spin, 
\begin{equation}
\sigma_{\mu}^{j}=\overbrace{\mathds{1}\otimes\mathds{1}\otimes\cdots\otimes\underbrace{\sigma_{\mu}}_{j^{th}\ \mbox{factor}}\otimes\cdots\otimes\mathds{1}}^{n\ \mbox{factors}}\;,\label{pauli3}
\end{equation}
with 
\begin{eqnarray}
\hspace*{-0.5cm}\sigma_{x}=\left(\begin{array}{cc}
0 & 1\\
1 & 0
\end{array}\right),\sigma_{y}=\left(\begin{array}{cc}
0 & -i\\
i & 0
\end{array}\right),\sigma_{z}=\left(\begin{array}{cc}
1 & 0\\
0 & -1
\end{array}\right),
\end{eqnarray}
and $\langle\sigma_{\mu}^{j}\rangle=\langle\Psi|\sigma_{\mu}^{j}|\Psi\rangle$
denotes the corresponding expectation value. The vectors ${\bf X}_{j}=(\langle\sigma_{x}^{j}\rangle,\langle\sigma_{y}^{j}\rangle,\langle\sigma_{y}^{j}\rangle)$,
${\bf X}_{j}\in{\mathbb{R}}^{3}$, allow a geometric representation
of each reduced state in ${\mathbb{R}}^{3}$, satisfying $0\le\|{\bf X}_{j}\|\le1$.
Since ${\sf tr}_{j}(\rho_{j}^{2})=\frac{1}{2}(1+\|{\bf X}_{j}\|^{2})$,
the state $\ket{\Psi}$ is entangled if $\|{\bf X}_{j}\|<1$ for at
least one $j$, represented by a point {\em inside} the corresponding
local Bloch sphere. One may therefore consider $\ket{\Psi}$ to be
maximally entangled if $\|{\bf X}_{j}\|=0$ for all $j$. On the other
hand, the state $\ket{\Psi}$ is unentangled (i.e., a product state)
if $\|{\bf X}_{j}\|=1$ for all $j$, corresponding to points lying
on the surface of the Bloch sphere.

@@

\subsection{Deutsch Algorithm}
